---
title: "Untitled"
author: "Rose Hartman"
date: "10/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("stargazer")
library("broom")
```

# Data import and tidying

NOTE: These are fake data. 

```{r data_import}
covid_testing <- read_csv("covid_testing.csv")

# test that it read in correctly
head(covid_testing)
```

Your tidying steps might be standardized in your lab, or you may be following the steps laid out in your pre-registration. Some data tidying will be the result of EDA (exploratory data analysis), when you discover anomolies 

```{r data_tidying}
# save a new version of the data for use in our models
# only keep test results that are either positive or negative (no invalid tests), and only for pediatric patients (< 18 years old)

model_data <- covid_testing %>% 
  # filter to only keep rows where result is either positive or negative
  filter(result %in% c("positive", "negative"),
         # age is less than 18
         age < 18) %>% 
  # calculate a total turn around time (collected to received plus received to verified)
  mutate(total_tat = col_rec_tat + rec_ver_tat) %>% 
  # also turn categorical variables into factors
  mutate(result = as.factor(result),
         gender = as.factor(gender),
         drive_thru_ind = factor(drive_thru_ind, levels = c(0,1), labels = c("no", "yes")),
         clinic_name = as.factor(clinic_name),
         demo_group = as.factor(demo_group)) %>% 
  # filter out extreme outliers, with TAT greater than 30 days, or less than 1 hour
  filter(total_tat < 30*24, 
         total_tat > 1)
```

# Models

## Hypothesis 1

Turn around time is shorter for drive-thru tests.

We can run this as a t-test, or as ordinary least squares regression (the models are equivalent).

```{r t_test}
hyp1_t <- t.test(total_tat ~ drive_thru_ind, data = model_data, var.equal = TRUE)
```

```{r ols_reg}
hyp1_lm <- lm(total_tat ~ drive_thru_ind, data = model_data)
```

```{r other_ols_examples, eval = FALSE}
?formula

# turn around time predicted from pandemic day
lm(total_tat ~ pan_day, data = model_data)

# same model as above, but allow for the possibility of a quadratic trend in time 
# see https://online.stat.psu.edu/stat462/node/158/
lm(total_tat ~ poly(pan_day, degree = 2), data = model_data)

# turn around time predicted from drive_thru, pan_day, and their interaction (i.e. perhaps the effect of pan_day is different for drive thru tests vs. non drive thru tests)
# see https://education.arcus.chop.edu/understanding-interactions/ 
lm(total_tat ~ drive_thru_ind*pan_day, data = model_data)
```


## Hypothesis 2

Probability of a positive test result increases with age, within the pediatric population.

Test with a logistic regression. Result (positive or negative) is the outcome, and age is the only predictor. 

```{r logistic_reg}
hyp2 <- glm(result ~ age, family = binomial, data = model_data)
```

# Examining model objects

```{r examine, eval=FALSE}
# for each saved model object, run these four lines of code in the console, one at a time
# substituting in the name you saved in place of "model_obj"
model_obj
summary(model_obj)
plot(model_obj)
str(model_obj)
```

You can use functions from the broom package to turn model objects into tidy tables, which makes them easier to work with.

```{r broom}
tidy(hyp1_lm) # a table of coefficients for the model
glance(hyp1_lm) # a table of model fit stats
augment(hyp1_lm) # returns your original data, but with things like model estimates and residuals added on as new columns
```

```{r resid_plots}
# a histogram of the model residuals
hyp1_lm %>% 
  augment() %>% 
  ggplot(mapping = aes(x=.resid)) + 
  geom_histogram()

# examining residuals separately for different values of drive_thru_ind
hyp1_lm %>% 
  augment() %>% 
  ggplot(mapping = aes(x=.resid)) + 
  geom_histogram() + 
  facet_wrap(~ drive_thru_ind)
```



# Presenting results

## Model tables


```{r, results='asis'}
stargazer(hyp1_lm, type = "html")

stargazer(hyp2, type = "html")
```

```{r, results='asis'}
# defaults
stargazer(hyp1_lm, hyp2, type = "html")

# some tweaking. see ?stargazer for more options
stargazer(hyp1_lm, hyp2, type = "html", 
          dep.var.labels = c("Turn Around Time", "Prob(Positive)"), 
          covariate.labels = c("Drive Thru", "Age"),
          model.numbers = FALSE)
```


## Inline stats

You can include R code and results within your text as well, which is particularly valuable when reporting stats. For example, let's say we want to report our logistic regression results. 

```{r}
z <- summary(hyp2)$coefficients[2,3]
p <- summary(hyp2)$coefficients[2,4]
```

A logistic regression predicting the probability of a positive result from patients' age revealed a significant relationship between age and result, such that older pediatric patients were more likely to have positive test results (Z = `r round(z, 2)`, p = `r round(p, 3)`).


```{r}
# a function to print p values
# 
# It prints "p = " and the exact p value rounded to three decimal places,
# except if p < .001, then it prints "p < .001"
report_p <- function(p){
  print_p <- ifelse(p < .001, "p < .001", 
                    ifelse(p >= .001, paste("p =", round(p, 3)), NA))
  return(print_p)
}
```

A logistic regression predicting the probability of a positive result from patients' age revealed a significant relationship between age and result, such that older pediatric patients were more likely to have positive test results (Z = `r round(z, 2)`, `r report_p(p)`).

