---
title: "Intro to Stats in R"
author: "Rose Hartman"
date: "10/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("stargazer")
library("broom")
```

# Data import and tidying

NOTE: These are fake data. 

```{r data_import}
covid_testing <- read_csv("covid_testing.csv")

# test that it read in correctly
head(covid_testing)
```

Your tidying steps might be standardized in your lab, or you may be following the steps laid out in your pre-registration. Some data tidying will be the result of EDA (exploratory data analysis), such as when you discover anomalies, outliers, or errors in the data. 

```{r data_tidying}
# save a new version of the data for use in our models
# only keep test results that are either positive or negative (no invalid tests), and only for pediatric patients (< 18 years old)

model_data <- covid_testing %>% 
  # filter to only keep rows where result is either positive or negative
  filter(result %in% c("positive", "negative"),
         # age is less than 18
         age < 18) %>% 
  # calculate a total turn around time (collected to received plus received to verified)
  mutate(total_tat = col_rec_tat + rec_ver_tat) %>% 
  # calculate a centered version of pan_day and age, if we want to use them as predictors in models with interactions
  mutate(pan_day_c = pan_day - mean(pan_day),
         age_c = age - mean(age)) %>% 
  # also turn categorical variables into factors
  mutate(result = as.factor(result),
         gender = as.factor(gender),
         drive_thru_ind = factor(drive_thru_ind, levels = c(0,1), labels = c("no", "yes")),
         clinic_name = as.factor(clinic_name),
         demo_group = as.factor(demo_group)) %>% 
  # filter out extreme outliers, with TAT greater than 30 days, or less than 1 hour
  filter(total_tat < 30*24, 
         total_tat > 1)
```

# Models

## Hypothesis 1

Turn around time is shorter for drive-thru tests.

We can run this as a t-test, or as ordinary least squares regression (the models are equivalent).

```{r t_test}
hyp1_t <- t.test( )
```

```{r ols_reg}
hyp1_lm <- lm( )
```


## Hypothesis 2

Probability of a positive test result increases with age, within the pediatric population.

Test with a logistic regression. Result (positive or negative) is the outcome, and age is the only predictor. 

```{r logistic_reg}
hyp2 <- glm( )
```

# Examining model objects

```{r examine, eval=FALSE}
# for each saved model object, run these four lines of code in the console, one at a time
# substituting in the name you saved in place of "model_obj"
model_obj
summary(model_obj)
plot(model_obj)
str(model_obj)
```

You can use functions from the broom package to turn model objects into tidy tables, which makes them easier to work with.

```{r broom, eval=FALSE}
tidy(model_obj) # a table of coefficients for the model
glance(model_obj) # a table of model fit stats
augment(model_obj) # returns your original data, but with things like model estimates and residuals added on as new columns
```

Some examples of how you might use the `augment` function to explore model results:

```{r resid_plots}
# a histogram of the model residuals


# examining residuals separately for different values of drive_thru_ind

```

```{r fitted_plot}
# plot predicted probability by age

```


# Presenting results

## Model tables


```{r, results='asis'}
# use stargazer to generate a model summary table for hyp1


# use stargazer to generate a model summary table for hyp2

```

```{r, results='asis'}
# put both models together in one table for comparison


```


## Inline stats

You can include R code and results within your text as well, which is particularly valuable when reporting stats. For example, let's say we want to report our logistic regression results. 

```{r}
z <- summary(hyp2)$coefficients[2,3]
p <- summary(hyp2)$coefficients[2,4]
```


